---
title: "experiment3: More Airline Data"
date: "2023-09-11"
output: html_notebook
---



```{r}
library(nycflights13)
library(tidyverse)
library(janitor)
library(here)
```



# Get data from the online source.

Save the download to your data_raw directory. It will be a zipfile, but that wil be okay. Change the name of the zip file to RawData.

## Load the BOT data

We take advantage of the here package to get the path to the file that we stored. We also ma note that inside the zipfile is a csv file. We will read the csvfile.

```{r}
thisfile=here("RawData","DL_SelectFields.zip")

df2022=read_csv(thisfile) %>% clean_names()
```

## Subsetting to data of interest

Let's focus on flights from La Guardia ("LGA") and eliminate cargo flights by requiring at least 1 passenger.

Call the resultant dataframe 'df9'.

```{r}
df9 = df2022 |> filter(passengers>0,origin=="LGA")
```

# Exploring the BOT 2022 data

## Subsetting to data of interest

Lets try to visualize a monthly count of flights using a barchart.

First, a not quite right visualization

```{r}
df9 |>
  ggplot(aes(month)) + geom_bar()
```
By default, 'geom_bar' is counting the number of rows, where we have asked it to visualize the count by 'month'.

** Take a look at the dataset and discover why counting rows is not going to give us a count of flights.**

the visualization we want is:

```{r}
df9 |> ggplot(aes(month)) + geom_bar(aes(weight=departures_performed))
```

Make some observations of this plot.

around the holidays (easter, thanksgiving, and christmas) the number of flights increased

## A new visualization

Can you make a boxplot where the bar height is based on the number of passengers riding each month.


```{r}
df9 |> ggplot(aes(month)) + geom_bar(aes(weight=passengers))
```

Observation:
* low passenger numbers in early 2022 likely due to COVID, but the number of flights were still up.


Here is a pretty but not so useful plot:

```{r}
df9 |>
  ggplot(aes(month,fill=carrier_name)) + geom_bar(aes(weight=departures_performed))
```

## Arrivals and Departures from LGA

```{r}
df10 = df2022 |> filter(passengers>0,origin=="LGA" | dest=="LGA")

df10 |> ggplot(aes(month)) + geom_bar(aes(weight=passengers))
```


```{r}
df11=df10 |> select(month,passengers,seats,carrier_name,dest,origin)
```


```{r}
df12 =df10 |> select(1:5, month,contains("id"))
```


```{r}
df13 = df11 |> mutate(percent_loading = passengers/seats*100)

df13 |> ggplot(aes(percent_loading)) + 
  geom_histogram()+facet_wrap(~carrier_name,scales="free_y")
```


## Class work on Wednesday 9/13/23

```{r}
#2017

thisfile=here("RawData", "BOT Airline Data-20230913","DL_SelectFields_2017.zip")

df2017=read_csv(thisfile) %>% clean_names()

#2018

thisfile=here("RawData", "BOT Airline Data-20230913","DL_SelectFields_2018.zip")

df2018=read_csv(thisfile) %>% clean_names()

#2019

thisfile=here("RawData", "BOT Airline Data-20230913","DL_SelectFields_2019.zip")

df2019=read_csv(thisfile) %>% clean_names()

#2020

thisfile=here("RawData", "BOT Airline Data-20230913","DL_SelectFields_2020.zip")

df2020=read_csv(thisfile) %>% clean_names()

#2021

thisfile=here("RawData", "BOT Airline Data-20230913","DL_SelectFields_2021.zip")

df2021=read_csv(thisfile) %>% clean_names()

#2022

thisfile=here("RawData", "BOT Airline Data-20230913","DL_SelectFields_2022.zip")

df2022=read_csv(thisfile) %>% clean_names()

#2023

thisfile=here("RawData", "BOT Airline Data-20230913","DL_SelectFields_2023.zip")

df2023=read_csv(thisfile) %>% clean_names()
```

### combine dataframes


```{r}
dfc=df2017 |>
  bind_rows(df2018) |>
  bind_rows(df2019) |>
  bind_rows(df2020) |>
  bind_rows(df2021) |>
  bind_rows(df2022) |>
  bind_rows(df2023) 

```


### arrivals and departures from LGA

```{r}
dfc1=dfc |>
  filter(passengers>0,origin=="LGA" | dest =="LGA")
```




```{r}
dfc1|> 
  ggplot(aes(month)) + 
  geom_bar(aes(weight=passengers)) +
  ggtitle("Total Passengers by Month over Multiyear Period")
```


```{r}
dfc1 |>
  mutate(month_year = year+month/12) |>
  ggplot(aes(month_year)) +
  geom_bar(aes(weight=passengers)) +
  ggtitle("Total Passengers by Month over Multiyear Period")
```


```{r}
dfc1 |> ggplot(aes(year)) +
  geom_bar(aes(weight=passengers)) +
  ggtitle("total passengers by year over multiyear period")
```

### Question - do more folks travel *LGA to MIA* or *MIA to LGA*


```{r}
dfc2=dfc1 |>
  filter(passengers>0, origin=="MIA" |dest == "MIA")

dfc2|> 
  mutate(month_year = year+month/12) |>
  ggplot(aes(month_year)) +
  geom_bar(aes(weight=passengers, color = origin), position = "dodge") +
  ggtitle("Total Passengers by Month over Multiyear Period")
```

### the bs homework of the day  -  try to break it down by years, represent different visualizations, 


